Reconstructing Propositional Proofs in Type Theory

### 2017-10-31 paper/sections/introduction.tex
[ paper ] minor fixes/ .

Previous studies have reported proof-reconstruction using
proof-assistants based on higher-order logic where the development is
at a mature
stage~\cite{paulson2010three,hurlin07practical,kaliszyk2013}. Another
approaches has been proposed for proof-reconstructing using type
theory~\cite{Bezem2002,armand2011,Ekici2017,kanso2016light}.

We describe a formal reconstruction of proofs generated by the \Metis
prover~\cite{hurd2003first}---our source system---in Martin-Löf type
theory~\cite{Nordstrom-Petersson-Smith-1990}. We
formalize the subset of the \Metis inference rules for the
propositional logic fragment using a syntactical treatment of the
logic. The \Metis reasoning was formalized in
\Agda~\cite{agdateam}---our target system---in two
libraries~\cite{AgdaProp,AgdaMetis} and we implemented a
proof-reconstruction tool named \Athena~\cite{Athena} written in
\Haskell that is able to generate \Agda proof-terms for \Metis
derivations.

@@

The source codes accompanying this paper (programs, libraries, and
examples) are available at \verb!GitHub!:
\begin{itemize}
  \item The \Athena program that translates proofs generated by \Metis to \Agda:
  \url{http://github.com/jonaprieto/athena}.
  \item The \verb!agda-prop! library as a formalization in \Agda for
    classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\item The \verb!agda-metis! library as a formalization in \Agda to
  justify \Metis proofs of classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\end{itemize}
The proof-reconstruction was carried on for \Metis~2.3
(release~20170822), and the programs and libraries were tested with
\Agda~2.5.3, \Agda standard library~0.14, and \prg{GHC}~8.2.1.
\end{document}

>>>>

Previous studies have reported proof-reconstruction using
proof-assistants based on higher-order logic where the development is
at a mature
stage~\cite{paulson2010three,hurlin07practical,kaliszyk2013}. Another
approaches has been proposed for proof-reconstructing based on type
theory~\cite{Bezem2002,kanso2016light,Keller2013}.

We describe a formal reconstruction of proofs generated by the \Metis
prover~\cite{hurd2003first}---our source system---in Martin-Löf type
theory~\cite{Nordstrom-Petersson-Smith-1990}. We
formalize the subset of the \Metis inference rules for the
propositional logic fragment using a syntactical treatment.
The \Metis reasoning was formalized in
\Agda~\cite{agdateam}---our target system---in two
libraries~\cite{AgdaProp,AgdaMetis} and we implemented a
proof-reconstruction tool named \Athena~\cite{Athena} written in
\Haskell that is able to generate \Agda proof-terms for \Metis
derivations.

@@

The source code accompanying this paper (programs, libraries, and
examples) is available at \verb!GitHub!:
\begin{itemize}
  \item The \Athena program that translates proofs generated by \Metis to \Agda:
  \url{http://github.com/jonaprieto/athena}.
  \item The \verb!agda-prop! library as a formalization in \Agda for
    classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\item The \verb!agda-metis! library as a formalization in \Agda to
  justify \Metis proofs of classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\end{itemize}
The proof-reconstruction was carried on for \Metis~2.3
(release~20170822), and the programs and libraries were tested with
\Agda~2.5.3, \Agda standard library~0.14, and \prg{GHC}~8.2.1.
\end{document}

### 2017-10-31 paper/sections/type-theory.tex
[ paper ] minor fixes/ .

Type theory is another formalism for the foundation of mathematics, and it has became in a key instrument to study logic and proof theory.
It was proposed in 1902 to avoid the Russell's paradox and it follows
the same philosophy of constructive mathematics where the witness of
a statement is \emph{everything}. In that direction, we could say
the main actors in type theory are the statements and their proofs.
By following the Curry-Howard correspondence (see, for example,~\cite{Wadler-2015}), we say
\emph{types} are theorems and their \emph{terms} are their proofs.

Since type theory is a formal system, we have a syntax and a set of
derivation rules. These rules enables us to derive a kind of conclusions called
\emph{judgments}. A judgment is another way to say a term has certain type.

@@

To reconstruct \Metis inference rules in type theory, we observed that
transformations performed by some rules are implemented by \emph{general
recursive} functions.

>>>>

Type theory is a formalism for the foundation of mathematics, and it has became in a key instrument to study logic and proof theory.
It was proposed in 1902 to avoid Russell's paradox and it follows
the same philosophy of constructive mathematics where the witness of
a statement is \emph{everything}. In that direction, we could say
the main actors in type theory are the statements and their proofs.
By following the Curry-Howard correspondence (see, for example,~\cite{Wadler-2015}), we say
\emph{types} are theorems and their \emph{terms} are their proofs.

Since type theory is a formal system, we have a syntax and a set of
derivation rules. These rules enables us to derive a kind of conclusions called
\emph{judgments}. A judgment is another way to say that a term has a certain type.

@@

To reconstruct \Metis inference rules in type theory, we observed that some
rules or their inner functions are implemented by \emph{general recursive}
functions.
